---
layout: default
title: Example with Multiple Data Types
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Multiple Data Types

In this example we will use the Bioconductor package
[regioneR](http://bioconductor.org/packages/release/bioc/html/regioneR.html) to create 
different sets of genomic regions and positions representing different data types: 
a small set of large regions, a number of point positions in the genome representing 
some kind of numeric value (expression, methylation...), a small number of positions with
labels to be used as markers and a set of overlapping medium regions to be plotted as
sequencing reads. In addition to [regioneR](http://bioconductor.org/packages/release/bioc/html/regioneR.html)
and [rakyoploteR](http://bioconductor.org/packages/release/bioc/html/karyoploteR.html) we 
will use the [zoo](https://cran.r-project.org/package=zoo) package to compute the rolling 
meand and sd of the data points.

## Data Creation

We will use different calls to **createRandomRegions** from regioneR to create the random 
positions and regions in the genome and calls to **runif** and **rnorm** to create the data
values:

```{r, message=FALSE, warning=FALSE}
library(karyoploteR)
library(regioneR)
library(zoo)

set.seed(1234)

#Parameters
  data.points.colors <- c("#FFBD07", "#00A6ED",  "#FF3366", "#8EA604", "#C200FB")

  num.data.points <- 3000
  num.big.regions.up <- 30
  num.big.regions.down <- 30

  num.mid.regions <- 6000
  
  num.marks <- 90

#Create the random fake data  

  #Big regions
  big.regs.up <- joinRegions(createRandomRegions(nregions = num.big.regions.up, length.mean = 20000000, length.sd = 10000000, non.overlapping = TRUE, mask=NA), min.dist = 1)
  big.regs.down <- joinRegions(createRandomRegions(nregions = num.big.regions.down, length.mean = 10000000, length.sd = 5000000, non.overlapping = TRUE, mask=big.regs.up), min.dist = 1)

  #Data points
  data.points <- createRandomRegions(nregions = num.data.points, length.mean = 1, length.sd = 0, non.overlapping = TRUE, mask=NA)
  mcols(data.points) <- data.frame(y=rnorm(n = num.data.points, 0.5, sd = 0.1))
  dp.colors <- sample(head(data.points.colors, 2), size = num.data.points, replace = TRUE)

  #and move the data points with the big regions
  data.points[overlapsAny(data.points, big.regs.up)]$y <- data.points[overlapsAny(data.points, big.regs.up)]$y + runif(n=numOverlaps(data.points, big.regs.up), min = 0.1, max=0.3)
  data.points[overlapsAny(data.points, big.regs.down)]$y <- data.points[overlapsAny(data.points, big.regs.down)]$y - runif(n=numOverlaps(data.points, big.regs.down), min = 0.1, max=0.3)

  #markers
  marks <- createRandomRegions(nregions = num.marks, length.mean = 1, length.sd = 0)
  mcols(marks) <- data.frame(labels=paste0("rs", floor(runif(num.marks, min = 10000, max=99999))))

  #medium regions
  mid.regs <- createRandomRegions(nregions = num.mid.regions, length.mean = 5000000, length.sd = 1000000)

```


## Plotting

Once we have all the data available we can start plotting it. We will create a Karyoplot of 
the human genome with 2 data panels, one over and one below the ideogram. In order to 
better see the data we will create a karyoplot with only 3 chromosomes.

### Data Panel 1

In the data panel 1, the top one, we will plot the big regions and the data points going from
the ideogram r0=0 to 80% of the panel r1=0.8 and we'll use the top 20% (r0=0.8, r1=1) to plot
the markers.

So we'll plot the big regions as rectangles using **kpRect**. We want them to span all the 
vertical space (y0=0, y1=1) in the first 80% of the panel (r0=0, r1=0.8) and to have no border
(border=NA). 

After that we will plot the data points with **kpPoints** in the exact same region as the 
rectangles. In addition, we will add an axis. 

Once the data points are plotted, we will add the mean and sd of the data points. We'll do 
that chromosome per chromosome since the **rollmean** and **rollaply** functions from zoo 
do not understand about chromosome separation. The mean will be plotted with **kpLines** 
and the sd around the mean with a call to **kpPlotRibbon**, that plots a variable width 
polygon from y0 to y1.

Finally, the markers will be plotted with a call to **kpPlotMarkers**, setting 
r1=1.1 so the tip the marker lines end slightly over the top of the big regions. 
With this function the marker labels will be moved as necessary to avoid label 
overlapping.


### Data Panel 2

In the data panel 2 we'll plot the medium regions in the 80% of the data panel furthest from 
the ideogram (r0=0.2, r1=1) and the coverage of these regions in the closest 20%.

To plot the regions we'll use the **kpPlotRegions** function, that automatically piles up
overlapping regions.

To plot the coverage we'll user the **kpPlotCoverage** that given a set of regions computes 
the coverage and plots it as an area plot. By default the coverage would have had the base
on the ideogram side, since all data panels are oriented from the ideogram out. Since we want
the coverage to be inverted, we can simply flip the r0 and r1 values (r0=0.2, r1=0) and the
plot will be flipped. This is a very useful trick to invert the plotting coordinates.

Finally, we add a horizontal bar depicting a fake 30x level with **kpAbline** and add a text
label outside the data panel margins with a simple **kpText**.

```{r Figure, message=FALSE, warning=FALSE,  fig.width=11, fig.height=11}
  kp <- plotKaryotype(plot.type = 2, chromosomes = c("chr1", "chr2", "chr3"))

  ### Data Panel 1 ###

  #Big regions
  kpRect(kp, data = big.regs.up, y0=0, y1=1, col="#FFDDDD", border=NA, r0=0, r1=0.8)
  kpRect(kp, data = big.regs.down, y0=0, y1=1, col="#DDFFDD", border=NA, r0=0, r1=0.8)
  
  #Data points
  kpAxis(kp, ymin = 0, ymax = 1, r0=0, r1=0.8, numticks = 5, col="#666666", cex=0.5)
  kpPoints(kp, data=data.points, pch=16, cex=0.5, col=dp.colors, r0=0, r1=0.8)
  
  #Mean and sd of the data points.  
  for(chr in seqlevels(kp$genome)) {
    chr.dp <- sort(keepSeqlevels(x = data.points, value = chr, pruning.mode = "coarse"))
    rmean <- rollmean(chr.dp$y, k = 6, align = "center")  
    rsd <- rollapply(data = chr.dp$y, FUN=sd, width=6)
    kpLines(kp, chr = chr, x=start(chr.dp)[3:(length(chr.dp)-3)], y=rmean, col=data.points.colors[3], r0=0, r1=0.8)
    kpPlotRibbon(kp, chr=chr, data=chr.dp[3:(length(chr.dp)-3)], y0=rmean-rsd, y1=rmean+rsd, r0=0, r1=0.8, col="#FF336633", border=NA)
  }
    
  #Markers
  kpPlotMarkers(kp, data=marks, label.color = "#333333", r1=1.1, cex=0.5, label.margin = 5)

  ### Data Panel 2 ###
    
  #medium regions and their coverage
    
    kpPlotRegions(kp, data = mid.regs, r0 = 0.2, r1=1, border=NA, data.panel=2)
    kpPlotCoverage(kp, data=mid.regs, r0=0.2, r1=0, col=data.points.colors[2], data.panel = 2)
    kpPlotCoverage(kp, data=mid.regs, r0=0.2, r1=0.12, col=data.points.colors[1], data.panel = 2)
    
    kpText(kp, chr=seqlevels(kp$genome), y=0.4, x=0, data.panel = 2, r0=0.2, r1=0, col="#444444", label="30x", cex=0.8, pos=2)
    kpAbline(kp, h=0.4, data.panel = 2, r0=0.2, r1=0, col=data.points.colors[3])
    

```



