---
layout: tutorial
label: PlotTypes
title: PlotTypes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 14)
knitr::opts_chunk$set(fig.height = 14)
set.seed(1234)
```

## Plot Types

karyoploteR has different plot types available, each one defining a global
layout of ideograms and one or more data panel around them. Plot types are 
specified at plot creation time in `plotKaryotype` with the `plot.type`
parameter and are identified by integers.

The simplest (and default) plot type is 1, wich represents the ideograms 
as horizontal bars vertically stacked with a single data panel above them. Plot 
type 2 adds a second data panel below the ideograms and plot types 3 to 5 
represent the chromosomes as a single line of horizontal ideograms (think a 
manhattan plot) with variation in the number and position of the data panels. 
Finally, plot types 6 and 7 do not have standard data panels and can be used to
create ideogram only plots.


These are the available data types:

### _plot.type=1_

```{r Figure1, message=FALSE, warning=FALSE}
library(karyoploteR)

kp <- plotKaryotype(plot.type=1, main="plot.type=1")
kpDataBackground(kp, data.panel = 1)
kpText(kp, chr="chr1", x=60e6, y=0.5, labels="data.panel=1", data.panel = 1)
```

### _plot.type=2_

```{r Figure2, message=FALSE, warning=FALSE}
kp <- plotKaryotype(plot.type=2, main="plot.type=2")
kpDataBackground(kp, data.panel = 1)
kpText(kp, chr="chr1", x=60e6, y=0.5, labels="data.panel=1", data.panel = 1)
kpDataBackground(kp, data.panel = 2)
kpText(kp, chr="chr1", x=60e6, y=0.5, labels="data.panel=2", data.panel = 2)
```

### _plot.type=3_

```{r Figure3, message=FALSE, warning=FALSE, fig.height=7}
kp <- plotKaryotype(plot.type=3, main="plot.type=3")
kpDataBackground(kp, data.panel = 1)
kpText(kp, chr="chr1", x=60e6, y=0.5, labels="data.panel=1", data.panel = 1)
kpDataBackground(kp, data.panel = 2)
kpText(kp, chr="chr1", x=60e6, y=0.5, labels="data.panel=2", data.panel = 2)
```

### _plot.type=4_

```{r Figure4, message=FALSE, warning=FALSE, fig.height=5}
kp <- plotKaryotype(plot.type=4, main="plot.type=4")
kpDataBackground(kp, data.panel = 1)
kpText(kp, chr="chr1", x=60e6, y=0.5, labels="data.panel=1", data.panel = 1)
```

### _plot.type=5_

```{r Figure5, message=FALSE, warning=FALSE, fig.height=5}
kp <- plotKaryotype(plot.type=5, main="plot.type=5")
kpDataBackground(kp, data.panel = 1)
kpText(kp, chr="chr1", x=60e6, y=0.5, labels="data.panel=1", data.panel = 1)
```

### _plot.type=6_

```{r Figure6, message=FALSE, warning=FALSE, fig.height=7}
kp <- plotKaryotype(plot.type=6, main="plot.type=6")
kpDataBackground(kp)
kpText(kp, chr="chr1", x=60e6, y=0.5, labels="data.panel=\"ideogram\"", data.panel = "ideogram")
```


### _plot.type=7_

**Note:** This plot type with full cytoband representation only makes sense for 
very wide plots or for a subset of chromosomes.

```{r Figure7, message=FALSE, warning=FALSE, fig.height=4, fig.width=18}
kp <- plotKaryotype(plot.type=7, main="plot.type=7")
kpDataBackground(kp, col="#CCCCCCCC")
kpText(kp, chr="chr1", x=60e6, y=0.5, labels="data.panel=\"ideogram\"", data.panel = "ideogram")
```


